---
title: "Assignment 4"
author: "Isabelle Radis, Laura Ingulsrud, Lauren Jack"
date: "11/9/2018"
output:
  html_document: default
  word_document: default
---

doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
if (doc.type == "docx") { pander(df) } else { kable(df) }

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo = FALSE, message = FALSE}

# Load necessary packages
library(tidyverse)
library(kableExtra)
library(vcdExtra)

```

```{r,echo = FALSE, message = FALSE}

# Read in .csv files
lob_size_abundance <- read.csv("lobster_size_abundance.csv")
lob_traps<- read.csv("lobster_traps.csv")

```

```{r,echo = FALSE, message = FALSE}

# Problem 1. Lobster abundance and fishing pressure (2012 - 2017). Describe trends in lobster abundance (counts) and fishing pressure (trap buoys) at the five locations from 2012 - 2017. Ignore transect information - we are only interested in evaluating abundance and pressure on the order of SITE. Note: you are not expected to use regression here - just think of ways to clearly describe annual totals visually and in text, noting important trends, events, and differences.


lob_abundance <- lob_size_abundance %>% # Create simple data frame of lobster abundance from lob_size_abundance data frame
  select(YEAR, SITE, COUNT) %>% # Select only YEAR, SITE, and COUNT columns
  filter(COUNT!=0) # Filter out all values of 0 in COUNT column

lob_traps_simple <- lob_traps %>% # Create simple data frame of lobster traps from lob_traps data frame
  select(YEAR, SITE, TRAPS) %>% # Select only YEAR, SITE, and TRAPS columns
  filter(SITE!="AHND",
         SITE!="ABUR",
         SITE!="AHND to AQUE",
         SITE!="GOLB") # Filter out all sites except the 5 LTER locations: AQUE, NAPL, MOHK, IVEE, and CARP

# Make scatter plot for mean annual lobster abundance by site (2012-2017)
abundance <- lob_abundance %>% 
 group_by(SITE,YEAR) %>% 
 summarize(total_count=round(sum(COUNT)), digits=1)%>% 
 ggplot(aes(x=YEAR,y= total_count)) +
 geom_point()+
 facet_wrap(~SITE, scale="free") +
 xlab("Year") +
 ylab("Total Count")
   
abundance # Call "abundance" scatter plot

# Make a column graph of total lobster abundance by site
abundance_col <- ggplot(lob_abundance, aes(x=YEAR, y=COUNT))+
  geom_col()+
  facet_wrap(~SITE, scale="free")

abundance_col # Call "abundance_col" column graph

# Make a line graph of total lobster abundance within each site (2012-2017)
abundance_line <- as.data.frame(lob_abundance) %>% 
  expand.dft(freq = "COUNT")%>% 
  count(YEAR, SITE) %>% 
  ggplot(aes(x=YEAR, y=n))+
  geom_line(aes(color=SITE))+
  xlab("Year")+
  ylab("Lobster Abundance (count)")+
  ggtitle("Abundance of Lobster from 2012-2017") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

abundance_line # Call "abundance_line" line graph

# Make scatter plot for annual traps set by site (2012-2017)
traps <- lob_traps_simple %>% 
 group_by(SITE,YEAR) %>% 
 summarize(total_count=round(sum(TRAPS)), digits=1)%>% 
 ggplot(aes(x=YEAR,y= total_count))+
 geom_point()+
 facet_wrap(~SITE, scale="free")+
 xlab("Year")+
 ylab("Total Traps")+
scale_y_continuous(expand = c(0,0)) 
   
traps # Call "traps" scatter plot

# Make a column graph of total traps set by site 
traps_col<- ggplot(lob_traps_simple, aes(x=YEAR, y=TRAPS))+
  geom_col()+
  facet_wrap(~SITE, scale="free")

traps_col # Call "traps_col" column graph

# Make a line graph of total traps at each site (2012-2017)
traps_line<- as.data.frame(lob_traps_simple) %>% 
  expand.dft(freq = "TRAPS") %>% 
  count(YEAR, SITE) %>% 
  ggplot(aes(x=YEAR, y=n))+
  geom_line(aes(color= SITE))+
  xlab("Year")+
  ylab("Fishing Pressure (Traps)")+
  ggtitle("Fishing Pressure from 2012-2017 at Non-MPA Sites") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black"))
 
traps_line # CAll "traps_line" line graph

```

```{r,echo = FALSE, message = FALSE}

# Problem 2. Compare mean lobster size by site in 2017. Compare mean lobster sizes (carapace length (mm)) across the five sites for lobster observations collected in 2017. Warning: the size data are not in tidy format. There are rows that contain size information for multiple lobsters observed (e.g., if the researcher saw 3 lobsters all with carapace length ~ 60 mm, then they will have a single row where COUNT = 3 and SIZE = 60). You’ll want to get this into case format - where each lobster has its own row - before doing statistical analyses. There are many ways to do this. One hint: function expand.dft in the vcdExtra package (it doesn’t like tibbles, so you might need to coerce to data.frame first). *Anova boxplot and error bars


# put lobster size data into case format
lob_abun_simple<-expand.dft(lob_size_abundance, freq="COUNT")

#sumarize abundance data
lab_simple<-lob_abun_simple %>% 
  select(SITE, SIZE) %>% 
  group_by(SITE) %>% 
  summarize(Mean= mean(SIZE),
           SD= sd(SIZE),
           Median= median(SIZE),
           Maximum= max(SIZE),
           SampleSize= length(SIZE))
           
lab_simple
# make collumn graph of mean lobster sizes  I THINK THIS IS VERY WRONG...TAKES TEH MEAN OF SIZES from ALL YEARS 
lab_col<-ggplot(lab_simple, aes(x= SITE, y = Mean))+
  geom_col(aes(fill= "pink"), color="white") +
  geom_errorbar(aes(ymin= Mean-SD, ymax= Mean+SD, width=.3))+
  xlab("SITE")+
  ylab("Mean size(mm)") +
  ggtitle("Mean Lobster Size by Site in 2017") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position= "none",
        axis.line = element_line(colour = "black"))+ 
  scale_y_continuous(expand = c(0,0)) +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  
lab_col
```

3. Changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes)
From the data description (http://sbc.lternet.edu/cgi-bin/showDataset.cgi?docid=knb-lter-sbc.77):
“Data on abundance, size and fishing pressure of California spiny lobster (Panulirus interruptus)
are collected along the mainland coast of the Santa Barbara Channel. Spiny lobsters are an
important predator in giant kelp forests off southern California. Two SBC LTER study reefs are
located in or near the California Fish and Game Network of Marine Protected Areas (MPA), Naples
and Isla Vista, both established as MPAs on 2012-01-01. MPAs provide a unique opportunity to
investigate the effects of fishing on kelp forest community dynamics. Sampling began in 2012 and
is ongoing.”
At Isla Vista and Naples Reef, the two protected MPA sites (with zero fishing pressure), how do lobster sizes
in 2012 and 2017 compare? At the non-MPA sites?

```{r,echo = FALSE, message = FALSE}

 lab_simple_12vs17<-lob_abun_simple %>% 
  select(YEAR, SITE, SIZE) %>% 
  filter(YEAR=="2017"| YEAR== "2012") %>% 
  group_by(SITE,YEAR) %>% 
  summarize(Mean= mean(SIZE),
           SD= sd(SIZE),
           Median= median(SIZE),
           Maximum= max(SIZE),
           SampleSize= length(SIZE))

lab_col_12vs17 <- ggplot(lab_simple_12vs17, aes(x= SITE, y = Mean, group= YEAR))+
  geom_col(aes(fill= YEAR), color="white",
           position="dodge") +
  xlab("SITE")+
  ylab("Mean size(mm)") +
  ggtitle("Mean Lobster Size by Site in 2012 and 2017") +
 scale_fill_gradient(c("gray70","gray40")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position= "none",
        axis.line = element_line(colour = "black"))+ 
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lab_col_12vs17
# add leggend with years...
# when I add the manual color, it says "continuous value supplied to discrete scale"

```

```{r,echo = FALSE, message = FALSE}

#Preform 5 two sample t-test to test if lobster sizes differ significantly between 2012 and 2016(one t test per site)

#H0: means are equal (differnce in means = 0)
#Ha: means are NOT equal (difference in means is NOT 0)


NAPL_abun_12<- lob_size_abundance %>% 
  filter(SITE=="NAPL", YEAR== "2012") %>% 
  pull("SIZE")
NAPL_abun_12

NAPL_abun_17<- lob_size_abundance %>% 
  filter(SITE=="NAPL", YEAR== "2017") %>% 
  pull("SIZE")
NAPL_abun_17

t_test<- t.test(NAPL_abun_17, NAPL_abun_12)
t_test

# p value is 2.842e^-14
# thus, there is a significant diference between the two means

# maybe a man whiney u
IVEE_abun_12<- lob_size_abundance %>% 
  filter(SITE=="IVEE", YEAR== "2012") %>% 
  pull("SIZE")
IVEE_abun_12

IVEE_abun_17<- lob_size_abundance %>% 
  filter(SITE=="IVEE", YEAR== "2017") %>% 
  pull("SIZE")
IVEE_abun_17

t_test<- t.test(IVEE_abun_17, IVEE_abun_12)
t_test 
# p value is 5.485e-08
# thus, there is a significant diference between the two means



AQUE_abun_12<- lob_size_abundance %>% 
  filter(SITE=="AQUE", YEAR== "2012") %>% 
  pull("SIZE")
AQUE_abun_12

AQUE_abun_17<- lob_size_abundance %>% 
  filter(SITE=="AQUE", YEAR== "2017") %>% 
  pull("SIZE")
AQUE_abun_17

t_test<- t.test(AQUE_abun_17, AQUE_abun_12)
t_test 
# p value = 0.7803
# thus, there is not a significant diference between the two means

CARP_abun_12<- lob_size_abundance %>% 
  filter(SITE=="CARP", YEAR== "2012") %>% 
  pull("SIZE")
CARP_abun_12

CARP_abun_17<- lob_size_abundance %>% 
  filter(SITE=="CARP", YEAR== "2017") %>% 
  pull("SIZE")
CARP_abun_17

t_test<- t.test(CARP_abun_17, CARP_abun_12)
t_test 
# p value = 4.158e^-16
# thus, there is a significant diference between the two means

MOHK_abun_12<- lob_size_abundance %>% 
  filter(SITE=="MOHK", YEAR== "2012") %>% 
  pull("SIZE")
MOHK_abun_12

MOHK_abun_17<- lob_size_abundance %>% 
  filter(SITE=="MOHK", YEAR== "2017") %>% 
  pull("SIZE")
MOHK_abun_17

MOHK_ftest<-var.test(MOHK_abun_17, MOHK_abun_12)
MOHK_ftest

t_test<- t.test(MOHK_abun_17, MOHK_abun_12, var.equal=TRUE)
t_test 

cohens.d
  
  
  
# p value = 0.1076
# thus, there is not a significant diference between the two means


#.. do this for all or do you lump NAPL and IVEE together to then say MPA sites have a significant difference (or dont) in lobster size from 2012 to 2017 
# then lump Non-MPA sites together and say there is (or is not) a significant difference in lobster size from 2012 to 2017


```


4. Proportions of “legal” lobsters at the 5 sites in 2017
The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site
are above the legal minimum? Does that proportion differ significantly across the 5 sites? Note: We’ll be
doing chi-square in labs next week, or go ahead with maximum resourcefulness and check out the chisq.test()
function on your own!

```{r}

```









